---
title: "Project 2"
author: "Symphony Hopkins"
date: "2023-03-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

**Read the dataset in Portland_Cement_Data.csv Download Portland_Cement_Data.csv into R. Name the loaded data Portland_Cement. And use head() function to show the first 6 rows.**

```{r, include = TRUE, echo = TRUE}
#importing dataset
Portland_Cement <- read.csv("~/Documents/Maryville_University/DSCI_618/Portland_Cement_Data.csv")

#displaying first 6 rows
head(x=Portland_Cement, n=6)
```

**Examine Data Graphically using strip charts, box plots, or other charts that can help you check if mixing techniques affect the strength of the cement.**

For this question, we will create a box plot.

```{r, include = TRUE, echo = TRUE}
#importing library
library(ggplot2)
#converting Mixing column to a factor
Portland_Cement$Mixing <- as.factor(Portland_Cement$Mixing)
#creating boxplot
ggplot(data=Portland_Cement, mapping=aes(x=Mixing, y=Strength)) + 
  geom_boxplot() + labs(title='Mixing Techniques vs Cement Strength')+
  stat_summary(geom='point', fun=mean)
```

From the box plots, it seems that the different mixing techniques affect the cement's strength; however, we will have to confirm this with additional testing.

**Test the hypothesis that mixing techniques affect the strength of the cement. Use α = 0.05**

```{r, include = TRUE, echo = TRUE}
#creating ANOVA model
Portland_Cement_AOV <- aov(Strength~Mixing, data = Portland_Cement)
summary(Portland_Cement_AOV)
```
Our p-value, 0.000489, is less than our significance level (α = 0.05), which means that we reject the null hypothesis and conclude that the treatment means differ.

**Construct a normal probability plot of the residuals (Normal Q-Q Plot). What conclusion would you draw about the validity of the Normality assumption?**

```{r, include = TRUE, echo = TRUE}
#creating Normal Q-Q Plot
plot(Portland_Cement_AOV, which=c(2))
```

For the most part, the data follows along the line, so the Normality assumption is correct.

**Plot the residuals versus the predicted tensile strength (Residuals vs Fitted). Comment on the plot.**

```{r, include = TRUE, echo = TRUE}
#creating Residuals vs Fitted Plot
plot(Portland_Cement_AOV, which=c(1))
```

The data points don't have any distinct patterns and they are evenly spread, which is a good indication that we don't have any non-linear relationships.

**Prepare a scatter plot of the residuals (Scale-Location plot). What conclusion would you draw about the validity of the Homoscedasticity assumption?**

```{r, include = TRUE, echo = TRUE}
plot(Portland_Cement_AOV, which=c(3))
```

We can conclude that the Homoscedasticity assumption is correct because the points are spread apart and there is a mostly horizontal line.

**Check if the following statements are true:**

* **The mean strength of Mixing Technique 1 differs significantly from Mixing Technique 3 at α = 0.05 level.**
* **The mean strength of Mixing Technique 2 differs significantly from Mixing Technique 4 at α = 0.05 level.**
* **The average of Mixing Technique 1 and 3 differs significantly from the mean strength of Technique 2 at α = 0.05 level.**

```{r, include = TRUE, echo = TRUE}

#assigning contrasts
tech_1_vs_tech_3 <- c(1,0,-1,0)
tech_2_vs_tech_4 <- c(0,1,0,-1)
tech_1_3_vs_tech_2 <- c(1,0,1,-2)
contrasts(Portland_Cement$Mixing) <- cbind(tech_1_vs_tech_3, tech_2_vs_tech_4,tech_1_3_vs_tech_2)
contrasts(Portland_Cement$Mixing)

#checking if mean strengths significantly differ
Portland_Cement_AOV_Contrasts <- aov(Strength ~ Mixing, data=Portland_Cement)
summary.lm(Portland_Cement_AOV_Contrasts)

```

From the results, we found the following:

* With a p-value of 0.650146, we conclude that the mean strength of Mixing Technique 1 *does not* differs significantly from Mixing Technique 3; therefore, we fail to reject the null hypothesis.
* With a p-value of 0.000636, we conclude that the mean strength of Mixing Technique 2 *does* differs significantly from Mixing Technique 4; therefore, we reject the null hypothesis.
* With a p-value of 0.481602, we conclude that the average of Mixing Technique 1 and 3 *does not * differ significantly from the mean strength of Technique 2; therefore, we fail to reject the null hypothesis.

**Calculate the Power of the analysis given the treatment means are μ1 = 2900 μ2 = 3100 μ3 = 2850 and μ4 = 2680, error variance = 230^2,and α = 0.05**

```{r, include = TRUE, echo = TRUE}
#calculating power of analysis
treatment_means <- c(2900,3100,2850,2680)
Portland_Cement_PA <- power.anova.test(groups=4, n=4, between.var=var(treatment_means), within.var=230^2, sig.level=.05)
Portland_Cement_PA
```

The results above show us that we have a 43.5% chance to get a significant result with the current settings.

